---
title: "BET 2026 Stepwise Guide"
subtitle: "Editable model folders, runtime settings, and Kflow launch checks"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    theme: cosmo
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
source("stepwise-config.R")

has_yaml <- requireNamespace("yaml", quietly = TRUE) && file.exists("kflow.yaml")
kflow_config <- if (has_yaml) yaml::read_yaml("kflow.yaml") else list()

as_text <- function(value, default = "") {
  if (is.null(value) || length(value) == 0 || is.na(value[[1]])) {
    return(default)
  }
  if (is.logical(value)) {
    return(tolower(as.character(value[[1]])))
  }
  as.character(value[[1]])
}

step_exists <- function(step_id) {
  dir.exists(file.path("steps", step_id))
}

model_dir <- function(step_id) {
  file.path("steps", step_id, "model")
}

source_status <- function(step_id, source_dir) {
  if (!is.na(source_dir) && nzchar(source_dir)) {
    if (identical(source_dir, ".")) {
      path <- file.path("steps", step_id)
    } else if (grepl("^/|^[A-Za-z]:", source_dir)) {
      path <- source_dir
    } else {
      path <- file.path("steps", step_id, source_dir)
    }
  } else {
    path <- model_dir(step_id)
  }

  if (dir.exists(path)) {
    "ok"
  } else {
    paste("missing:", path)
  }
}

model_checks <- within(stepwise_models, {
  step_folder <- ifelse(vapply(step_id, step_exists, logical(1)), "ok", "missing")
  source_folder <- mapply(source_status, step_id, source_dir, USE.NAMES = FALSE)
  run_mode_valid <- ifelse(run_mode %in% c("last_par", "single", "doitall"), "ok", "check")
})
```

# What To Edit

Most edits should happen in only two places:

- `stepwise-config.R`: choose default settings and register each model row.
- `steps/<step_id>/model/`: keep the MFCL input files for that model.

Use `patch.R` inside a step folder only when a model needs a small scripted edit before MFCL runs. Each step folder is independent, so a new sensitivity or review model should usually be a new numbered folder under `steps/`.

# Current Defaults

```{r defaults-table}
defaults <- data.frame(
  setting = c(
    "default_step_select",
    "flow_group",
    "trigger_next",
    "mfcl_fevals",
    "docker_image",
    "program_path"
  ),
  value = c(
    stepwise_value("default_step_select", "01-base-11par"),
    stepwise_value("flow_group", "bet-2026-e2e"),
    stepwise_value("trigger_next", "true"),
    stepwise_value("mfcl_fevals", ""),
    as_text(kflow_config$docker_image, "ghcr.io/pacificcommunity/tuna-flow:v1.5"),
    as_text(kflow_config$env$PROGRAM_PATH, "/home/mfcl/mfclo64")
  ),
  meaning = c(
    "Model selection used when STEP_SELECT is not supplied.",
    "Kflow group label used to connect stepwise, plot, and report jobs.",
    "Whether command-line Kflow submissions keep the downstream plot/report chain.",
    "Blank uses the row-level fevals value; a number overrides selected rows.",
    "Docker image used by Kflow and local Docker runs.",
    "MFCL executable path inside the Docker image."
  ),
  stringsAsFactors = FALSE
)
knitr::kable(defaults)
```

# Model Rows

```{r model-table}
model_display <- stepwise_models[, c(
  "step_id",
  "enabled",
  "model_label",
  "run_mode",
  "input_par",
  "frq",
  "output_par",
  "fevals",
  "notes"
)]
knitr::kable(model_display)
```

# Folder Checks

```{r checks-table}
checks_display <- model_checks[, c(
  "step_id",
  "enabled",
  "run_mode",
  "step_folder",
  "source_folder",
  "run_mode_valid"
)]
knitr::kable(checks_display)
```

# Run Modes

`last_par` continues from a numbered `.par` file and writes the next numbered file. For example, `11.par` becomes `12.par`. This is the usual quick model-check mode.

`single` runs exactly the listed input and output `.par` files. Use it when the output name should be explicit instead of inferred.

`doitall` runs `doitall.sh` from the model folder. `MFCL_FEVALS` is only passed as an environment variable in this mode; the script must read it explicitly if a quick test should use it.

# Launch Examples

```bash
make list
make local STEP_SELECT=01-base-11par PROGRAM_PATH=/path/to/mfclo64
make docker STEP_SELECT=01-base-11par
make kflow STEP_SELECT=01-base-11par
make kflow STEP_SELECT=01-base-11par TRIGGER_NEXT=false
```

Use comma-separated model IDs to run several independent folders in one job:

```bash
make docker STEP_SELECT=01-base-11par,03-review-11par
make kflow STEP_SELECT=01-base-11par,03-review-11par
```

# Adding A Model

1. Copy an existing folder under `steps/`.
2. Rename it with the next numbered ID, such as `04-steepness-low`.
3. Put the model's MFCL inputs in `steps/04-steepness-low/model/`.
4. Copy one row in `stepwise-config.R` and update `step_id`, `model_label`, `run_mode`, and notes.
5. Run `make list`, then `make docs`, then launch with `STEP_SELECT=04-steepness-low`.

# Downstream Jobs

Kflow stores the normal dependency chain as `stepwise -> plot -> report`. The `stepwise_run$trigger_next` value controls the default behavior for command-line submissions through `make kflow`.

Set `TRIGGER_NEXT=false` for a one-off model-only run:

```bash
make kflow STEP_SELECT=01-base-11par TRIGGER_NEXT=false
```

Keep `TRIGGER_NEXT=true` when the selected model should feed the plot and report tasks after the stepwise job succeeds.
